---
title: "Pressure pain algometry reliability using the variance components approach"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

# Load packages

```{r message=FALSE, warning=FALSE}
# Helper
library(tidyverse)
library(boot)
library(parallel)
library(foreach)
# Stats
library(lme4)
library(psych)
library (hemp)
# Markdown
library(knitr)
#Plots
library (cowplot)
```

# Read in data

```{r, include = FALSE}
DaveData <- read_delim("DaveData.txt", "\t",
                       escape_double = FALSE, trim_ws = TRUE)
```

```{r}
load ("ICCdat.Rdata")
```


# Convert data to long format

```{r}
ppt.df = DaveData %>%
  na.omit() %>%
  gather (key = Var, value = PPT, c(-c(1:4))) %>%
  separate(Var, c("side_session", "reps"), sep = "_") %>%
  mutate (side = str_remove (side_session, "[0-9]+"),
          session = str_remove(side_session, "L|R")) %>%
  mutate (reps = as.numeric(reps))  %>%
  select (subj, side, session, reps, PPT) %>%
  mutate (subj = factor (subj),
          side = factor (side),
          session  = factor (session))
```

# Plotting 

```{r}

#tiff("../Manuscript/Plot2.tiff", width = 4, height = 4, units = 'in', res = 300, compression = 'none')

ppt.df %>% 
  mutate (session = recode (session,
                            "1" = "session 1",
                            "2" = "session 2"),
          side = recode (side,
                         "L" = "Left",
                         "R" = "Right")) %>%
  group_by (side, 
           session, 
           reps) %>%
  summarize (Mean = mean (PPT),
             Sd = sd (PPT)) %>%
  ungroup() %>%
  ggplot (aes (x = reps, y = Mean)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin =Mean - Sd,  ymax=Mean+Sd)) +
  facet_wrap (side~session, nrow =2) +
  ylab ("PPT (N)") +
  xlab ("Repetitions") +
  scale_x_continuous(breaks=seq(2,10,2)) +
  theme_cowplot()

#dev.off ()
```

# Method: Variance components approach

## Cross nested design

```{r}
f1 <- PPT ~ (1|subj) + (1|subj:session) + (1|subj:side) + (1|subj:session:side)

m1 <- lmer (f1, data = ppt.df)

varcomp1 = as.data.frame (VarCorr(m1))
varcomp1[,c(4,5)] <- round (varcomp1[,c(4,5)], 2)
varcomp1

round (confint(m1, oldNames=FALSE)^2, 2)
```

## Gstudy

```{r, include = FALSE}
g1 <- hemp::gstudy(m1)
g1$gstudy.out[,2] <- round (g1$gstudy.out[,2], 2)
g1
```

## Get ICC from variance components with bootstrap to get confidence interval

https://github.com/kohleth/kohleth.info/blob/master/content/post/2017-10-09-all-the-different-iccs.Rmd

### Create custom ICC function

`what` is what ICC does one want to calculate for? Example: `what` = `"reps"` calculates ICC(reps).

`k` is the number of repeated measurements for ICCk. 

```{r}
icc_type = function (what, k) {
  
  function (m) {
    
    what <- paste0(what, collapse = "|")
    
    v <- as.data.frame(VarCorr(m)) [c(1,4)]
    v[nrow (v), 1] <- "reps"
    n.mat <- matrix(nrow = nrow(v), ncol = length(k))
    for (i in 1:length(k)) n.mat[grep(names(k)[i], v$grp), 
                                i] <- k[i]
    n.mat[nrow(n.mat), ] <- k
    v$n <- apply(n.mat, 1, prod, na.rm = T)
    v$vcov.n <- v$vcov/v$n

    # ICC2 --------------------------------------------
    ## Total variance
    denom <- sum (v$vcov)
    ## Variance without what
    numer <- sum (v[!grepl (what, v$grp), "vcov"])
    ## ICC
    ICC2 <- numer/ denom
    
    #ICC2k---------------------------------------------
    ## Total variance with k repeats
    if (min(k) == 1) k <- NULL
    
    denom_k <- sum (v$vcov.n)
    numer_k <- sum (v[!grepl (paste0(c (what, names (k)), collapse = "|"), v$grp), "vcov"])
    ## ICC
    ICC2k <- numer_k / denom_k
    
    icc = c("ICC2" = ICC2, "ICC2k" = ICC2k)
    
    return (icc)
    
  }

}

```


### Create ICC boot function

```{r}

icc_boot <- function (m.mer, my_icc, B){
  
  ICCs = ICCboot = NULL  

  ICCs = my_icc(m.mer)

  ICCboot =list ( ICC2 = bootMer(m.mer, my_icc, nsim = B))
  
  calcBootCI=function(b,ind=1){
    boot.ci(b,type="perc",index=ind)$percent[4:5]
  }
  
  out = list (ICC2 = calcBootCI(ICCboot$ICC2,1),
              ICC2k = calcBootCI(ICCboot$ICC2,2)) %>%
    do.call (rbind, .)
  
  out=data.frame(ICCs[names (ICCs)],out)
  colnames(out)=c("val","lower","upper")
  
  return (out)
}

```

### Practice booting

```{r, eval = FALSE}

icc_reps <- icc_type (what = "session", k = c ("side" = 2, "reps" = 5))

icc_boot (m.mer = m1,
          B = 10, 
          my_icc = icc_reps)

```

## ICC(session)

```{r, eval = FALSE}

icc_reps <- icc_type (what = "session", k = c ("side" = 1, "reps" = 1))

icc_sess <- icc_boot (m.mer = m1,
                      B = 1000, 
                      my_icc = icc_reps)
```
```{r}
icc_sess
```

## ICC(session, reps)

```{r, eval = FALSE}

icc_reps <- icc_type (what = c("session", "reps"), k = c ("side" = 1, "reps" = 1))

icc_sess_rep <-icc_boot (m.mer = m1,
                    B = 1000, 
                    my_icc = icc_reps)
```
```{r}
icc_sess_rep 
```


## ICC(session, side, reps)

```{r,eval = FALSE}

icc_reps <- icc_type (what = c("session", "reps", "side"), k = c ( "reps" = 1))

icc_sess_side_rep <-icc_boot (m.mer = m1,
                    B = 1000, 
                    my_icc = icc_reps)
```
```{r}
icc_sess_side_rep
```

## ICCk(session, reps)

```{r, eval = FALSE}

doParallel::registerDoParallel(7)

ntimes <- 10
nboot <- 1000

boot.list = foreach (n = 1:ntimes, .packages = c("tidyverse", "lme4", "boot")) %dopar%{
  icc_reps <- icc_type (what = c("session", "reps"), k = c ("side" = 2, "reps" = n))
  
  icc_boot (m.mer = m1, 
          B = nboot, 
          my_icc = icc_reps)
}

names (boot.list) <- 1:ntimes

boot.df = bind_rows(boot.list, .id = "Reps") %>%
  rownames_to_column(var = "ICC") %>%
  mutate (ICC = ifelse (grepl ("ICC2k", ICC), "ICC2K", "ICC2"),
          Reps = as.numeric (Reps)) %>%
  filter (ICC == "ICC2K" & Reps > 1)
```

### Plot boot results

```{r}
#tiff("../Manuscript/Plot3.tiff", width = 4, height = 4, units = 'in', res = 300, compression = 'none')

ggplot (data = boot.df , aes (x = Reps, y = val)) +
  geom_line (position=position_dodge(1)) +
  geom_point(position=position_dodge(1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(1)) +
  labs (y =  "ICCk(session)",
         x = "Repetitions" ) + 
  scale_x_continuous(breaks=seq(2,10,2)) + 
  geom_hline(yintercept=0.8, linetype="dashed", 
                color = "black", size=1) + 
  theme_cowplot()

#dev.off()


```

# Save data

```{r, eval = FALSE}
save (DaveData, 
      ppt.df, 
      icc_sess,
      icc_sess_rep,
      icc_sess_side_rep,
      boot.list, 
      boot.df, 
      file = "ICCdat.Rdata")


```