---
title: "Pressure pain algometry reliability using the variance components approach"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

# Load packages

```{r message=FALSE, warning=FALSE}
# Helper
library(tidyverse)
library(boot)
library(parallel)
library(foreach)
# Stats
library(lme4)
library(psych)
library (hemp)
# Markdown
library(knitr)
#Plots
library (cowplot)
```

# Read in data

```{r message=FALSE, warning=FALSE}
DaveData <- read_delim("DaveData.txt", "\t",
                       escape_double = FALSE, trim_ws = TRUE)
```

# Convert data to long format

```{r}
ppt.df = DaveData %>%
  na.omit() %>%
  gather (key = Var, value = PPT, c(-c(1:4))) %>%
  separate(Var, c("side_session", "reps"), sep = "_") %>%
  mutate (side = str_remove (side_session, "[0-9]+"),
          session = str_remove(side_session, "L|R")) %>%
  mutate (reps = as.numeric(reps))  %>%
  select (subj, side, session, reps, PPT) %>%
  mutate (subj = factor (subj),
          side = factor (side),
          session  = factor (session))
```

# Plotting 

```{r}

tiff("../Manuscript/Plot2.tiff", width = 4, height = 4, units = 'in', res = 300, compression = 'none')

ppt.df %>% 
  mutate (session = recode (session,
                            "1" = "session 1",
                            "2" = "session 2"),
          side = recode (side,
                         "L" = "Left",
                         "R" = "Right")) %>%
  group_by (side, 
           session, 
           reps) %>%
  summarize (Mean = mean (PPT),
             Sd = sd (PPT)) %>%
  ungroup() %>%
  ggplot (aes (x = reps, y = Mean)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin =Mean - Sd,  ymax=Mean+Sd)) +
  facet_wrap (side~session, nrow =2) +
  ylab ("PPT (N)") +
  xlab ("Repetitions") +
  scale_x_continuous(breaks=seq(2,10,2)) +
  theme_cowplot()

dev.off ()
```

# Method: Variance components approach

## All factors crossed

Some evidence this model is too complex.

```{r}
f1 = PPT ~ (1|subj) + (1|session) + (1|side) + (1|reps) +
  (1|subj:session) + (1|subj:side) + (1|subj:reps) + (1|session:side) + (1|session:reps) + (1|side:reps) + 
  (1|subj:session:side) + (1|subj:session:reps) + (1|subj:side:reps) + (1|session:side:reps) 

m1 = lmer (f1, data = ppt.df)

varcomp1 = as.data.frame (VarCorr(m1))
varcomp1[,c(4,5)] <- round (varcomp1[,c(4,5)], 2)
varcomp1
```

## Gstudy

```{r}
g1 <- hemp::gstudy(m1)
g1
```


## Merge side and session to become separate trials

```{r}

ppt.df.merge <- ppt.df %>%
  unite ("trials", session, side, sep =  "_")

f2 = PPT ~ (1|subj) + (1|trials) + (1|reps) + (1|subj:reps) + (1|trials:reps) + (1|subj:trials)
m2 = lmer (f2, data = ppt.df.merge, REML = TRUE)

varcomp2 = as.data.frame (VarCorr(m2))
varcomp2[,c(4,5)] <- round (varcomp2[,c(4,5)], 2)
varcomp2

```

## Gstudy

```{r}
g2 <- hemp::gstudy(m2)
g2

```

## Get ICC from variance components with bootstrap to get confidence interval

https://github.com/kohleth/kohleth.info/blob/master/content/post/2017-10-09-all-the-different-iccs.Rmd

### Create custom ICC function

`what` is what ICC does one want to calculate for? Example: `what` = `"reps"` calculates ICC(reps).

`k` is the number of repeated measurements for ICCk. 

```{r}
icc_type = function (what, k) {
  
  function (m) {
    
    v=as.data.frame(VarCorr(m))
    vc = v$vcov
    name_4_res = paste0 (v$grp[!grepl(":|Residual", v$grp)], collapse = ":")
    names (vc)= v$grp
    names(vc) [names(vc) == "Residual"] = name_4_res

    # ICC2 --------------------------------------------
    ## Total variance
    denom = sum (vc)
    ## Variance without what
    numer = sum (vc[!grepl (what, names (vc))])
    ## ICC
    ICC2 = numer/ denom
    
    #ICC2k---------------------------------------------
    ## Total variance with k repeats
    denom_k = sum (vc[grepl (what, names (vc))]/ k) + sum (vc[!grepl (what, names (vc))])
    ## ICC
    ICC2k = numer/ denom_k
    
    return (icc = c("ICC2" = ICC2, "ICC2k" = ICC2k))
    
  }

}

```


### Create ICC boot function

```{r}

icc_boot <- function (formula, data, m.mer, my_icc, B){
  
  ICCs = ICCboot = NULL  

  ICCs = my_icc(m.mer)

  ICCboot =list ( ICC2 = bootMer(m.mer, my_icc, nsim = B))
  
  calcBootCI=function(b,ind=1){
    boot.ci(b,type="perc",index=ind)$percent[4:5]
  }
  
  out = list (ICC2 = calcBootCI(ICCboot$ICC2,1),
              ICC2k = calcBootCI(ICCboot$ICC2,2)) %>%
    do.call (rbind, .)
  
  out=data.frame(ICCs[names (ICCs)],out)
  colnames(out)=c("val","lower","upper")
  
  return (out)
}

```

### Practice booting

```{r}

icc_reps <- icc_type (what = "reps", k = 10)

icc_boot (formula = f2, 
          m.mer = m1,
          B = 10, 
          my_icc = icc_reps)

```

### How many repetitions do I need to perform to get my ICC > 0.8

```{r}

doParallel::registerDoParallel(7)

ntimes <- 10
B = 500

boot.list = foreach (n = 1:ntimes, .packages = c("tidyverse", "lme4", "boot")) %dopar%{
  icc_reps <- icc_type (what = "reps", k = n)
  
  icc_boot (formula = f2, 
          data = ppt.df.merge, 
          B = 20, 
          my_icc = icc_reps)
}

names (boot.list) <- 1:ntimes

boot.df = bind_rows(boot.list, .id = "Reps") %>%
  rownames_to_column(var = "ICC") %>%
  mutate (ICC = ifelse (grepl ("ICC2k", ICC), "ICC2K", "ICC2"),
          Reps = as.numeric (Reps))
```

### Plot boot results

```{r}
tiff("Plot4.tiff", width = 4, height = 4, units = 'in', res = 300, compression = 'none')

ggplot (data = boot.df , aes (x = Reps, y = val)) +
  geom_line (position=position_dodge(1)) +
  geom_point(position=position_dodge(1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(1)) +
  facet_wrap(~ICC) + 
  labs (y =  expression (paste("var ",Delta, "PPT (",N^2, ")")) ,
         x = "Repetitions" ) + 
  geom_hline(yintercept=0.8, linetype="dashed", 
                color = "black", size=1) + 
  theme_cowplot()

dev.off()


```

# Method: Traditional ICC (taking the average of N repetitions) 

```{r message=FALSE, warning=FALSE}
icc.list.m1 = list()

for (n in 1:10){
  # average
  icc.df.m1 = ppt.df %>%
  spread (key = session, value = PPT) %>%
  rename (sess1 = "1",
          sess2 = "2") %>%
  filter ( reps <= n) %>%
  group_by(subj) %>%
  summarize (sess1 = mean (sess1),
             sess2 = mean (sess2)) %>%
  ungroup () %>%
  as.data.frame()
  icc.list.m1 [[n]] = ICC (x = icc.df.m1[, -c(1)]) [[1]][c(2,5), c(1,2,7,8)]

}

icc.sf.ave = bind_rows(icc.list.m1 , .id = "Reps") %>%
  select (Reps, type, everything ()) %>%
  mutate (Reps = as.numeric (Reps))

colnames (icc.sf.ave) [4:5] = c("lower", "upper")

tiff("Plot3.tiff", width = 7, height = 4, units = 'in', res = 300, compression = 'none')

ggplot (data = icc.sf.ave , aes (x = Reps, y = ICC)) +
  geom_line (position=position_dodge(1)) +
  geom_point(position=position_dodge(1)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.2,
                 position=position_dodge(1)) +
  facet_wrap( ~ type, ncol = 2) +
  ylab ("ICC (Session)") +
  xlab ("Repetitions") +
  scale_colour_manual(values = c("red", "blue")) +
  geom_hline(yintercept=0.8, linetype="dashed", 
                color = "black", size=1) +
  scale_x_continuous(breaks=seq(0,10,2)) +
  theme_cowplot ()

dev.off ()

icc.sf %>%
  filter (Reps ==1 | Reps == 10) %>%
  mutate (ICC = round (ICC, 2),
          lower = round (lower, 2),
          upper = round (upper, 2))
```

# Save data

```{r, eval = FALSE}
save (DaveData, 
      ppt.df, 
      boot.list_sess1,
      boot.list_sess2 , 
      icc.list.m1, 
      icc.list.m2, 
      boot.list, 
      file = "ICCdat.Rdata")


```